Supporting Information:\\ Headgroup structure and cation binding in phosphatidylserine lipid bilayers

### 2019-05-22 Manuscript/manuscriptPSsuppl.tex
Update simulation details text

\noindent
{\it POPC:POPS (5:1) mixtures with additional calcium} \\
POPC:POPS (5:1) mixtures containing total 300 (250:50) lipids
%were simulated with the additional calcium concentrations.
%The data is available from Refs. \citenum{POPC5POPS1withCaClCHARMM} and \citenum{POPC5POPS1with1MCaClCHARMM}.
and 0.26 M and 1 M CaCl$_2$ were prepared using the CHARMM-GUI \cite{lee16,jo18}.
Systems were simulated using Gromacs 5 \cite{abraham2015gromacs} and CHARMM36 force field with the simulation parameters given
by the CHARMM-GUI at 298 K.

\subsection{CHARMM36ua}
\noindent
{\it DOPS and POPS bilayers.} \\
%The data in Table \ref{PSsystems} and Refs \citenum{charmm36uaDOPS303K,charmm36uaPOPS298K}
Starting structures for CHARMM36-UA DOPS and POPS simulations were taken from those generated for the CHARMM36 simulations (see above),
with any extraneous hydrogen atoms removed in the lipid tails. Force field parameters for DOPS and POPS were constructed through a
combination of the published CHARMM36 PS~\cite{venable13} and CHARMM36-UA PC parameters~\cite{lee14}.
% [REF: https://pubs.acs.org/doi/10.1021/jp410344g].
All other simulation parameters and conditions were identical to those described for the CHARMM36 all-atom DOPS and POPS simulations.

@@

\noindent
{\it POPS bilayers.}
%The data in Table \ref{PSsystems} and Refs \citenum{macrogPOPS298Kcorrect,macrogPOPS298KwithK}
%\todo{T. Piggot and M. Javanainen, please write simulation details} \\
%\hl{Piggot: please write about system with Na$^+$ counterions}.
Force field parameters for POPS simulations with the MacRog OPLS-AA force field~\cite{maciejewski14,kulig15b,rog16}
were constructed manually using the provided PS parameters as a guide~\cite{rog16}. The original parameters were
not used due to an incorrect arrangement of the {\it sn}-1 and {\it sn}-2 tails (i.e. the lipid was OPPS and not POPS).
The starting structure was the same as that used in the CHARMM36 POPS simulations (with Na+ ions) and was converted into
the correct format using a script to adjust the atom order. Water was treated using the TIP3P model~\cite{jorgensen83}.
Simulation parameters were chosen to closely mimic those used in the original force field publications~\cite{maciejewski14,kulig15b,rog16}:
van der Waals interactions were truncated at 1.0 nm with a dispersion correction applied to the energy and pressure;
Coulombic interactions were truncated at 1.0 nm with PME used to treat interactions beyond this cut-off.
All other simulation settings were identical to those described above for the CHARMM36 DOPS and POPS simulations,
including use of the Verlet cut-off scheme~\cite{Pall13}, apart from the use of LINCS to constrain all bonds during these simulations.

This membrane was further simulated after replacing Na$^+$ counterions by K$^+$ ones, which 
were described by the \AA{}qvist parameters \cite{aqvist90}. Other topologies
were unchanged. The simulation parameters were the same as used for the POPC/POPS mixtures with 
the MacRog force field, see below. The simulation was run for 200~ns, out of which 150~ns was used 
in the analyses.

@@

\subsection{Lipid17}
{\it DOPS and POPS bilayers.} \\

The structure of the bilayers with 128 DOPS or 128 POPS lipids were
obtained from CHARMM-GUI~\cite{lee16,jo18}. Bilayers are solvated with
4480 water molecules resulting in 31.1 water molecules per lipid. For
each system, 128 NaCl molecules are added to neutralize the system.
In all simulations, TIP3P~\cite{jorgensen1983comparison} was used as the solvent. SETTLE~\cite{miyamoto92} algorithm was employed 
to constrain the bonds in TIP3P water molecules. Simulations were ran using Amber18 simulation package \cite{amber18md}
and lipid parameters were obtained from Amber Lipid17 force field \cite{gould18}. For the ions, one set of simulations were run with Joung-Cheatham
ion parameters~\cite{joung2008determination} and one set of simulations were run with ff99SB (\AA{}qvist)~\cite{aqvist90} ion parameters.
The data in Table~\ref{PSsystems} and Refs~\citenum{lipid17POPSjcions,lipid17POPSff99ions,lipid17DOPSjcions,lipid17DOPSff99ions} contain
the number of molecules in simulations, related simulation files and trajectories.

Each bilayer structure was first minimized for 2500 steps with steepest
descent and an additional 2500 step with the conjugent gradient
algorithms while restrainig the solvent and ion molecules with a force
constant 500~kcal~mol$^{-1}$~\AA{}$^{-1}$. An additional minimization
step was used with the same parameters after removing the constraints on
the solvent and ion molecules. During the minimization procedure,
non-periodic boundary conditions were applied, no pressure scaling was
used and none of the bonds were restrained. A three step heating
procedure was applied: in the first heating step the system temperature
is increased from 0.0~K to 100.0~K using Langevin thermostat with
5.0~ps$^{-1}$ collision frequency while restraining the lipid positions
with harmonic springs of force constant
20.0~kcal~mol$^{-1}$~\AA{}$^{-1}$ at constant volume without periodic
boundary conditions. In the second heating step, the system temperature
was increased from 100.0~K to 200.0~K using Langevin thermostat with
5.0~ps$^{-1}$ collision frequency within 10000 steps while restraining
the lipid positions with harmonic springs of force constant
10.0~kcal~mol$^{-1}$~\AA{}$^{-1}$ at constant volume. In the third
heating step, the system temperature is increased from 200.0~K to
303.0~K within 10000 steps using Langevin thermostat with 5.0~ps$^{-1}$
collision frequency while restraining the lipid positions with harmonic
springs of force constant 10.0~kcal~mol$^{-1}$~\AA{}$^{-1}$ at 1~bar
constant pressure using Berendsen barostat with 3.0~ps relaxation time.
From the second heating step onward, periodic boundary conditions in all
dimensions were applied. The semi-isotropic pressure coupling along
\textit{xy}-plane was employed and the surface tension was set to
0.0~dyn explicitely. After the third heating step, each structure is
equilibriated for 100~ns at 1~bar and 303~K constant pressure using
Berendsen barostat and temperature using Langevin thermostat,
respectively. For the equilibriation step, semi-isotropic pressure
coupling along \textit{xy}-plane was employed and the surface tension
was set to 0.0~dyn explicitely. For all heating and equilibration steps,
the lengths of the hydrogen containing bonds were restrained using SHAKE
algorithm and a time step of 2~fs was used. In all steps with periodic
boundary conditions, the non-bonded interactions are calculated with
particle mesh Ewald~\cite{essman95} using 1.0~nm cutoff.\\

For the production runs, the temperature and pressure were set to 303~K
and 1~bar, using Langevin thermostat with 1.0~ps$^{-1}$ collision
frequency and Berendsen barostat with 1.0~ps relaxation time,
respectively. Periodic boundary conditions were applied in all
dimensions and the non-bonded interactions are calculated with particle
mesh Ewald~\cite{essman95} using 1.0~nm cutoff. The lengths of the
hydrogen containing bonds were constrained using SHAKE algorithm and a
time step of 2~fs was used. Trajectories are saved in every 10~ps
intervals. Two independent trajectories of 500~ns long was generated and
the last 200~ns portions were used for the analysis.\\

\noindent
{\it POPC:POPS (5:1) mixtures with additional potassium and sodium ions} \\

The initial structure of the bilayers with 120 POPC and 24 POPS lipids were obtained from CHARMM-GUI~\cite{lee16,jo18}.
Bilayers were solvated with 5760 TIP3P water molecules~\cite{jorgensen1983comparison}. For the simulations of the lipids and ions,
the Amber Lipid 17 \cite{gould18} and ff99SB~\cite{aqvist90} force fields were used, respectively.
Simulations were ran using Amber18 simulation package \cite{amber18md} 
The simulation steps are the same as the the simulations of
DOPS and POPS bilayers described above, with the only difference that the total simulation time was 300~ns, of which last 200~ns
was used for the analysis. The number of ions in each specific system is given in Table~\ref{mixedIONsystemsMONOVALENT}.
The trajectory files can be found in Refs~\citenum{POPCpopsLIPID17withKCI,POPCpopsLIPID17withK,POPCpopsLIPID17withNaCI,POPCpopsLIPID17withNa}.\\

\noindent
{\it POPC:POPS (5:1) mixtures with additional calcium chloride concentration} \\
%The data in Table \ref{mixedIONsystems} and Refs \citenum{lipid17_cacl_series,POPCpopsLIPID17withCaCl} 
We have performed two sets of simulations of POPC:POPS (5:1) mixtures using Amber Lipid17 force field \cite{gould18}
with additional calcium chloride concentrations, 
which employed different simulation engines and parameters for ions.

One set of simulations has been performed using Amber18 simulation package \cite{amber18md} and ff99SB~\cite{aqvist90} ion parameters. 
Number of ions and water molecules for this set of simulations are shown in Table~\ref{mixedIONsystems}. The simulation details, minimization, heating, and equilibration runs are the same as described above for the other POPC:POPS (5:1) simulations with Amber Lipid17 forcefield. 
For each ion concentration, 100~ns equilibration runs were followed by 300~ns production runs with two independent trajectories, of which last 200~ns were used for the analysis. The trajectory files can be found in Ref~\citenum{POPCpopsLIPID17withCaCl}

Another set of simulations with Lipid17 force field for lipids \cite{gould18}
and Dang parameters for ions \cite{smith94, dang06}
was performed using the GROMACS 2018 simulation package~\cite{abraham2015gromacs}. 
We used the time step of 2~fs. 
The non-bonded interactions were calculated directly within 1.0~nm cutoff using the Verlet scheme \cite{Pall13}, 
and the long-range electrostatic forces were calculated using particle mesh Ewald~\cite{essman95}. 
The bond lengths of hydrogen atoms were constrained using LINCS algorithm \cite{hess97}.
Temperature was coupled to the velocity rescaling thermostat \cite{bussi07} at 298~K with a coupling constant of 1~ps.
Pressure was coupled to the Parrinello-Rahman barostat \cite{parrinello81} at 1~bar with a coupling constant of 10~ps. 
The composition of the individual simulations is shown in Table~\ref{mixedIONsystems}. 
Initial structures obtained from CHARMM-GUI~\cite{lee16,jo18} 
were equilibrated prior to production simulations used for analysis,
which are available in Ref.~\citenum{lipid17_cacl_series} including the simulation settings. 
Topologies in a format for the GROMACS package are available in the repository \citenum{ECCpops}.

\subsection{Slipids}
%
%\noindent
%{\it POPS}
%\todo{POPS \cite{slipidsPOPS298K} to be written by Piggot}%%
%
%\noindent
%{\it DOPS with 128 lipids.}
%\todo{DOPS with 128 lipids \cite{slipidsDOPS303K} to be written by Piggot}
%
\noindent
{\it POPS and DOPS with 128 lipids}
Starting structures for the Slipids POPS and DOPS (with 128 lipids) simulations were the same as constructed
for the CHARMM36 simulations. Force field parameters were taken from the published Slipids PS model,
however simulation cut-offs used herein were different to these previously published simulations~\cite{jambeck2012another}.
% [REF: https://pubs.acs.org/doi/10.1021/ct300777p].
This was due to the use of the older GROMACS group-based cut-off scheme in these previous simulations.
In the simulations reported in this work, the Verlet cut-off scheme was used. Van der Waals interactions were
truncated at 1.0 nm with LJ-PME~\cite{wennberg15}
% [REF: https://pubs.acs.org/doi/10.1021/acs.jctc.5b00726]
applied to account for the long-range interactions.
Coulombic interactions were truncated at 1.0 nm with PME used for the long-range interactions.
Full testing and validation of these settings with the Slipids PS force field will be provided in an additional forthcoming work.
All other simulation settings were identical to those described above for the CHARMM36 DOPS and POPS simulations.

\noindent
{\it DOPS with 288 lipids.}
The starting structure with 288 DOPS lipids, 11232 water molecules and 288 Na ions
was contructed with the MEMBRANE BUILDER website~\cite{ghahremanpour13}.
The Slipids~\cite{jambeck12,jambeck2012another} forcefield was used for DOPS, the TIP3P~\cite{jorgensen83} water model was used to solvate the system and
ions are described by the parameters derived by \AA{}qvist~\cite{aqvist90}.
The system was simulated in NPT ensemble for 200~ns using the GROMACS 5.0.4 package~\cite{abraham2015gromacs},
and the last 100~ns was used for the analysis.
Timestep of 2~fs was used with the leapfrog integrator. The Nos\'{e}--Hoover thermostat~\cite{nose84,hoover85} was used with reference temperature of 303~K and a relaxation time constant of 0.5~ps; lipids and water plus ions were coupled separately to the heat bath. Pressure was kept constant at 1.013~bar using a semi--isotropic Parrinello--Rahman
barostat~\cite{parrinello81} with a time constant of 10.0~ps. Long-range electrostatic interactions were calculated using the PME method~\cite{darden93,essman95}. A real space cut-off of 1.0~nm was employed with grid spacing of 0.12~nm in the reciprocal space. Lennard-Jones potentials were cut off at 1.4~nm, with a dispersion correction applied to both energy and pressure. All covalent bonds in lipids were constrained using the LINCS algorithm~\cite{hess97}, whereas water molecules were constrained using SETTLE~\cite{miyamoto92}. Twin-range cutoffs, 1.0~nm and 1.6~nm, were used for the neighbor lists with the long-range neighbor list updated every 10 steps.

@@

\noindent
{\it DOPS and POPS bilayers.} \\
%The data in Table \ref{PSsystems} and Refs \citenum{bergerPOPS298K,bergerDOPS303K}
%\todo{T. Piggot, please write simulation details} \\
Starting structures for DOPS and POPS simulations performed with a Berger-based PS force field
were taken from those produced for the CHARMM36 simulations. A script was used to re-order the
atoms and to remove any extraneous hydrogen atoms. Force field parameters were taken from those
published by Mukhopadhyay et al.~\cite{mukhopadhyay04}
% [REF: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1303994/]
and the SPC water model was used~\cite{berendsen81}. The simulation cut-offs closely matched
those of Mukhopadhyay et al.: van der Waals interactions were truncated at 1.0 nm;
Coulombic interactions were truncated at 1.0 nm with PME applied to account for the interactions beyond the cut-off.
All other simulation settings were identical to those described above for the CHARMM36 DOPS and POPS simulations,
including the use of the Verlet cut-off scheme, apart from all bonds were constrained using LINCS in these simulations.

@@

\noindent
{\it DOPS and POPS bilayers.} 
%The data in Table \ref{PSsystems} and Refs. \citenum{ckp1POPS303K,ckp2POPS303K,ckp1DOPS303K,ckp2DOPS303K}
%\todo{T. Piggot, please write simulation details} \\

Starting structures for GROMOS-CKP and GROMOS-CKPM simulations were taken from those produced for the CHARMM36 simulations.
A script was used to re-order the atoms and to remove any extraneous hydrogen atoms. The force fields parameters for DOPS and POPS
were as discussed above. The SPC water model was used in these simulations~\cite{berendsen81}. Simulation cut-offs were chosen
to closely match those used in the original GROMOS-CKP force field validation: van der Waals interactions were truncated at 1.4 nm
with a dispersion correction applied to the energy and pressure; Coulombic interactions were also truncated at 1.4 nm with PME used
to treat the long-range interactions. All other simulation settings were identical to those described above for the CHARMM36 DOPS
and POPS simulations bar, including the use of the Verlet cut-off scheme, apart from all bonds were constrained using LINCS in these simulations.

\noindent
{\it POPC:POPS (5:1) mixtures without additional ions}
% The data in Table \ref{mixedIONsystems} and Refs. \citenum{POPCpopsGROMOSCKPwithNa,POPCpopsGROMOSCKPMwithNa}
% \todo{T. Piggot, please write simulation details} \\

Starting structures for GROMOS-CKP and GROMOS-CKPM simulations were taken from those produced for the equivalent CHARMM36 simulations (see above).
A script was used to re-order the atoms and to remove any extraneous hydrogen atoms. Simulation setting were identical to those as described above
for the GROMOS-CKP(M) POPS bilayers.

>>>>

\noindent
{\it POPC:POPS (5:1) mixtures with additional calcium} 
POPC:POPS (5:1) mixtures containing total 300 (250:50) lipids
%were simulated with the additional calcium concentrations.
%The data is available from Refs. \citenum{POPC5POPS1withCaClCHARMM} and \citenum{POPC5POPS1with1MCaClCHARMM}.
and 0.26 M and 1 M CaCl$_2$ were prepared using the CHARMM-GUI \cite{lee16,jo18}.
Systems were simulated using Gromacs 5 \cite{abraham2015gromacs} and CHARMM36 force field with the simulation parameters given
by the CHARMM-GUI at 298 K.

\subsection{CHARMM36ua}
\noindent
{\it DOPS and POPS bilayers.} 
%The data in Table \ref{PSsystems} and Refs \citenum{charmm36uaDOPS303K,charmm36uaPOPS298K}
Starting structures for CHARMM36-UA DOPS and POPS simulations were taken from those generated for the CHARMM36 simulations (see above),
with any extraneous hydrogen atoms removed in the lipid tails. Force field parameters for DOPS and POPS were constructed through a
combination of the published CHARMM36 PS~\cite{venable13} and CHARMM36-UA PC parameters~\cite{lee14}.
% [REF: https://pubs.acs.org/doi/10.1021/jp410344g].
All other simulation parameters and conditions were identical to those described for the CHARMM36 all-atom DOPS and POPS simulations.

@@

\noindent
{\it POPS bilayers with Na$^+$ counterions.}
%The data in Table \ref{PSsystems} and Refs \citenum{macrogPOPS298Kcorrect,macrogPOPS298KwithK}
%\todo{T. Piggot and M. Javanainen, please write simulation details} \\
%\hl{Piggot: please write about system with Na$^+$ counterions}.
Force field parameters for POPS simulations with the MacRog OPLS-AA force field~\cite{maciejewski14,kulig15b,rog16}
were constructed manually using the provided PS parameters as a guide~\cite{rog16}. The original parameters were
not used due to an incorrect arrangement of the {\it sn}-1 and {\it sn}-2 tails (i.e. the lipid was OPPS and not POPS).
The corrected parameters are available from Ref.~\citenum{macrogPOPS298KwithK}.
The starting structure was the same as that used in the CHARMM36 POPS simulations (with Na+ ions) and was converted into
the correct format using a script to adjust the atom order. Water was treated using the TIP3P model~\cite{jorgensen83}.
Simulation parameters were chosen to closely mimic those used in the original force field publications~\cite{maciejewski14,kulig15b,rog16}:
van der Waals interactions were truncated at 1.0 nm with a dispersion correction applied to the energy and pressure;
Coulombic interactions were truncated at 1.0 nm with PME used to treat interactions beyond this cut-off.
All other simulation settings were identical to those described above for the CHARMM36 DOPS and POPS simulations,
including use of the Verlet cut-off scheme~\cite{Pall13}, apart from the use of LINCS to constrain all bonds during these simulations.

\noindent
{\it POPS bilayers with K$^+$ counterions.}
The membrane with sodium counterions was further simulated after replacing Na$^+$ counterions by K$^+$ ones, which 
were described by the \AA{}qvist parameters \cite{aqvist90}. Other topologies
were unchanged. The simulation parameters were the same as used for the POPC/POPS mixtures with 
the MacRog force field, see below. The simulation was run for 200~ns, out of which 150~ns was used 
in the analyses.

@@

\subsection{Lipid17}
{\it DOPS and POPS bilayers.} 
The structure of the bilayers with 128 DOPS or 128 POPS lipids were
obtained from CHARMM-GUI~\cite{lee16,jo18}. Bilayers are solvated with
4480 water molecules resulting in 31.1 water molecules per lipid. For
each system, 128 NaCl molecules are added to neutralize the system.
In all simulations, TIP3P~\cite{jorgensen1983comparison} was used as the solvent. SETTLE~\cite{miyamoto92} algorithm was employed 
to constrain the bonds in TIP3P water molecules. Simulations were ran using Amber18 simulation package \cite{amber18md}
and lipid parameters were obtained from Amber Lipid17 force field \cite{gould18}. For the ions, one set of simulations were run with Joung-Cheatham
ion parameters~\cite{joung2008determination} and one set of simulations were run with ff99SB (\AA{}qvist)~\cite{aqvist90} ion parameters.
The data in Table~\ref{PSsystems} and Refs~\citenum{lipid17POPSjcions,lipid17POPSff99ions,lipid17DOPSjcions,lipid17DOPSff99ions} contain
the number of molecules in simulations, related simulation files and trajectories.

Each bilayer structure was first minimized for 2500 steps with steepest
descent and an additional 2500 step with the conjugent gradient
algorithms while restrainig the solvent and ion molecules with a force
constant 500~kcal~mol$^{-1}$~\AA{}$^{-1}$. An additional minimization
step was used with the same parameters after removing the constraints on
the solvent and ion molecules. During the minimization procedure,
non-periodic boundary conditions were applied, no pressure scaling was
used and none of the bonds were restrained. A three step heating
procedure was applied: in the first heating step the system temperature
is increased from 0.0~K to 100.0~K using Langevin thermostat with
5.0~ps$^{-1}$ collision frequency while restraining the lipid positions
with harmonic springs of force constant
20.0~kcal~mol$^{-1}$~\AA{}$^{-1}$ at constant volume without periodic
boundary conditions. In the second heating step, the system temperature
was increased from 100.0~K to 200.0~K using Langevin thermostat with
5.0~ps$^{-1}$ collision frequency within 10000 steps while restraining
the lipid positions with harmonic springs of force constant
10.0~kcal~mol$^{-1}$~\AA{}$^{-1}$ at constant volume. In the third
heating step, the system temperature is increased from 200.0~K to
303.0~K within 10000 steps using Langevin thermostat with 5.0~ps$^{-1}$
collision frequency while restraining the lipid positions with harmonic
springs of force constant 10.0~kcal~mol$^{-1}$~\AA{}$^{-1}$ at 1~bar
constant pressure using Berendsen barostat with 3.0~ps relaxation time.
From the second heating step onward, periodic boundary conditions in all
dimensions were applied. The semi-isotropic pressure coupling along
\textit{xy}-plane was employed and the surface tension was set to
0.0~dyn explicitely. After the third heating step, each structure is
equilibriated for 100~ns at 1~bar and 303~K constant pressure using
Berendsen barostat and temperature using Langevin thermostat,
respectively. For the equilibriation step, semi-isotropic pressure
coupling along \textit{xy}-plane was employed and the surface tension
was set to 0.0~dyn explicitely. For all heating and equilibration steps,
the lengths of the hydrogen containing bonds were restrained using SHAKE
algorithm and a time step of 2~fs was used. In all steps with periodic
boundary conditions, the non-bonded interactions are calculated with
particle mesh Ewald~\cite{essman95} using 1.0~nm cutoff.

For the production runs, the temperature and pressure were set to 303~K
and 1~bar, using Langevin thermostat with 1.0~ps$^{-1}$ collision
frequency and Berendsen barostat with 1.0~ps relaxation time,
respectively. Periodic boundary conditions were applied in all
dimensions and the non-bonded interactions are calculated with particle
mesh Ewald~\cite{essman95} using 1.0~nm cutoff. The lengths of the
hydrogen containing bonds were constrained using SHAKE algorithm and a
time step of 2~fs was used. Trajectories are saved in every 10~ps
intervals. Two independent trajectories of 500~ns long was generated and
the last 200~ns portions were used for the analysis.

\noindent
{\it POPC:POPS (5:1) mixtures with additional potassium and sodium ions} 
The initial structure of the bilayers with 120 POPC and 24 POPS lipids were obtained from CHARMM-GUI~\cite{lee16,jo18}.
Bilayers were solvated with 5760 TIP3P water molecules~\cite{jorgensen1983comparison}. For the simulations of the lipids and ions,
the Amber Lipid 17 \cite{gould18} and ff99SB~\cite{aqvist90} force fields were used, respectively.
Simulations were ran using Amber18 simulation package \cite{amber18md} 
The simulation steps are the same as the the simulations of
DOPS and POPS bilayers described above, with the only difference that the total simulation time was 300~ns, of which last 200~ns
was used for the analysis. The number of ions in each specific system is given in Table~\ref{mixedIONsystemsMONOVALENT}.
The trajectory files can be found in Refs~\citenum{POPCpopsLIPID17withKCI,POPCpopsLIPID17withK,POPCpopsLIPID17withNaCI,POPCpopsLIPID17withNa}.

\noindent
{\it POPC:POPS (5:1) mixtures with additional calcium chloride concentration with default Amber parameters} 
%The data in Table \ref{mixedIONsystems} and Refs \citenum{lipid17_cacl_series,POPCpopsLIPID17withCaCl} 
%We have performed two sets of simulations of POPC:POPS (5:1) mixtures using Amber Lipid17 force field \cite{gould18}
%with additional calcium chloride concentrations, 
%which employed different simulation engines and parameters for ions.
Simulations were performed using Amber18 simulation package \cite{amber18md} with ff99SB~\cite{aqvist90} ion parameters and Amber Lipid17~\cite{gould18} force field. 
Number of ions and water molecules for this set of simulations are shown in Table~\ref{mixedIONsystems}.
The simulation details, minimization, heating, and equilibration runs are the same as described above for the other POPC:POPS (5:1) simulations with Amber Lipid17 forcefield. 
For each ion concentration, 100~ns equilibration runs were followed by 300~ns production runs with two independent trajectories, of which last 200~ns were used for the analysis.
The trajectory files can be found from Ref~\citenum{POPCpopsLIPID17withCaCl}.

\noindent
{\it POPC:POPS (5:1) mixtures with additional calcium chloride concentration with Dang ion parameters} 
The Lipid17 parameters for POPS were obtained from AmberTools18 \cite{amber18md} 
and converted to Gromacs format using acpype \cite{silva12}.  
The previously generated \cite{botan15} Lipid14 parameters for POPC in Gromacs format were downloaded from Ref.~\citenum{lipid14files}. 
Dang parameters for ions \cite{smith94, dang06} were used.
% Another set of simulations with Lipid17 force field for lipids \cite{gould18}
Initial structures were generated using CHARMM-GUI~\cite{lee16,jo18} 
and simulations were performed using the GROMACS 2018 simulation package~\cite{abraham2015gromacs}
with the time step of 2~fs. 
The non-bonded interactions were calculated directly within 1.0~nm cutoff using the Verlet scheme \cite{Pall13}, 
and the long-range electrostatic forces were calculated using particle mesh Ewald~\cite{essman95}. 
The bond lengths of hydrogen atoms were constrained using LINCS algorithm \cite{hess97}.
Temperature was coupled to the velocity rescaling thermostat \cite{bussi07} at 298~K with a coupling constant of 1~ps.
Pressure was coupled to the Parrinello-Rahman barostat \cite{parrinello81} at 1~bar with a coupling constant of 10~ps. 
The compositions of individual simulations are shown in Table~\ref{mixedIONsystems}. 
The systems were equilibrated 50~ns prior to production simulations used for analysis,
which are available in Ref.~\citenum{lipid17_cacl_series} including the simulation settings. 
Topologies in a format for the GROMACS package are available in the repository \citenum{ECCpops}.

\subsection{Slipids}
%
%\noindent
%{\it POPS}
%\todo{POPS \cite{slipidsPOPS298K} to be written by Piggot}%%
%
%\noindent
%{\it DOPS with 128 lipids.}
%\todo{DOPS with 128 lipids \cite{slipidsDOPS303K} to be written by Piggot}
%
\noindent
{\it POPS and DOPS with 128 lipids}
Starting structures for the Slipids POPS and DOPS (with 128 lipids) simulations were the same as constructed
for the CHARMM36 simulations. Force field parameters were taken from the published Slipids PS model~\cite{jambeck13},
however simulation cut-offs used herein were different to these previously published simulations~\cite{jambeck2012another}.
% [REF: https://pubs.acs.org/doi/10.1021/ct300777p].
This was due to the use of the older GROMACS group-based cut-off scheme in these previous simulations.
In the simulations reported in this work, the Verlet cut-off scheme was used. Van der Waals interactions were
truncated at 1.0 nm with LJ-PME~\cite{wennberg15}
% [REF: https://pubs.acs.org/doi/10.1021/acs.jctc.5b00726]
applied to account for the long-range interactions.
Coulombic interactions were truncated at 1.0 nm with PME used for the long-range interactions.
Full testing and validation of these settings with the Slipids PS force field will be provided in an additional forthcoming work.
All other simulation settings were identical to those described above for the CHARMM36 DOPS and POPS simulations.

\noindent
{\it DOPS with 288 lipids.}
The starting structure with 288 DOPS lipids, 11232 water molecules and 288 Na$^+$ ions
was contructed with the MEMBRANE BUILDER website~\cite{ghahremanpour13}.
The Slipids~\cite{jambeck13} forcefield was used for DOPS, the TIP3P~\cite{jorgensen83} water model was used to solvate the system and
ions are described by the parameters derived by \AA{}qvist~\cite{aqvist90}.
The system was simulated in NPT ensemble for 200~ns using the GROMACS 5.0.4 package~\cite{abraham2015gromacs},
and the last 100~ns was used for the analysis.
Timestep of 2~fs was used with the leapfrog integrator. The Nos\'{e}--Hoover thermostat~\cite{nose84,hoover85} was used with reference temperature of 303~K and a relaxation time constant of 0.5~ps; lipids and water plus ions were coupled separately to the heat bath. Pressure was kept constant at 1.013~bar using a semi--isotropic Parrinello--Rahman
barostat~\cite{parrinello81} with a time constant of 10.0~ps. Long-range electrostatic interactions were calculated using the PME method~\cite{darden93,essman95}. A real space cut-off of 1.0~nm was employed with grid spacing of 0.12~nm in the reciprocal space. Lennard-Jones potentials were cut off at 1.4~nm, with a dispersion correction applied to both energy and pressure. All covalent bonds in lipids were constrained using the LINCS algorithm~\cite{hess97}, whereas water molecules were constrained using SETTLE~\cite{miyamoto92}. Twin-range cutoffs, 1.0~nm and 1.6~nm, were used for the neighbor lists with the long-range neighbor list updated every 10 steps.

@@

\noindent
{\it DOPS and POPS bilayers.} 
%The data in Table \ref{PSsystems} and Refs \citenum{bergerPOPS298K,bergerDOPS303K}
%\todo{T. Piggot, please write simulation details} \\
Starting structures for DOPS and POPS simulations performed with a Berger-based PS force field
were taken from those produced for the CHARMM36 simulations. A script was used to re-order the
atoms and to remove any extraneous hydrogen atoms. Force field parameters were taken from those
published by Mukhopadhyay et al.~\cite{mukhopadhyay04}
% [REF: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1303994/]
and the SPC water model was used~\cite{berendsen81}. The simulation cut-offs closely matched
those of Mukhopadhyay et al.: van der Waals interactions were truncated at 1.0 nm;
Coulombic interactions were truncated at 1.0 nm with PME applied to account for the interactions beyond the cut-off.
All other simulation settings were identical to those described above for the CHARMM36 DOPS and POPS simulations,
including the use of the Verlet cut-off scheme, apart from all bonds were constrained using LINCS in these simulations.

@@

\noindent
{\it DOPS and POPS bilayers.} 
%The data in Table \ref{PSsystems} and Refs. \citenum{ckp1POPS303K,ckp2POPS303K,ckp1DOPS303K,ckp2DOPS303K}
%\todo{T. Piggot, please write simulation details} \\
Starting structures for GROMOS-CKP and GROMOS-CKPM simulations were taken from those produced for the CHARMM36 simulations.
A script was used to re-order the atoms and to remove any extraneous hydrogen atoms. The force fields parameters for DOPS and POPS
were as discussed above. The SPC water model was used in these simulations~\cite{berendsen81}. Simulation cut-offs were chosen
to closely match those used in the original GROMOS-CKP force field validation: van der Waals interactions were truncated at 1.4 nm
with a dispersion correction applied to the energy and pressure; Coulombic interactions were also truncated at 1.4 nm with PME used
to treat the long-range interactions. All other simulation settings were identical to those described above for the CHARMM36 DOPS
and POPS simulations, including the use of the Verlet cut-off scheme, apart from all bonds were constrained using LINCS in these simulations.

\noindent
{\it POPC:POPS (5:1) mixtures without additional ions}
% The data in Table \ref{mixedIONsystems} and Refs. \citenum{POPCpopsGROMOSCKPwithNa,POPCpopsGROMOSCKPMwithNa}
% \todo{T. Piggot, please write simulation details} \\
Starting structures for GROMOS-CKP and GROMOS-CKPM simulations were taken from those produced for the equivalent CHARMM36 simulations (see above).
A script was used to re-order the atoms and to remove any extraneous hydrogen atoms. Simulation setting were identical to those as described above
for the GROMOS-CKP(M) POPS bilayers.
